---
title: 'Lab 7: advanced plotting'
author: "Isaias Perez"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{R libraries}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(ggpubr)
library(plotly)

options(scipen = 999)
```

***1.** To begin, one of the aspects of programming I want to emphasize
is oftentimes you can re-use old code in order to save you time and
frustration. More specifically, as you gain more experience coding you
should find yourself regularly going back to older code you’ve written
to find ways to execute a task you’ve done previously without having to
start from scratch. From there, you can just modify your code as needed
so that you aren’t regularly re-learning how to do something.*

***a.** For question 1, save mtcars as data and then repurpose our code
from Lab 5 that turns our rownames into columns to create a column
called “Model” within ‘data’ that contains the car models. Then use the
head() function (used in Lab 1) to show me the first 5 rows of your new
dataset.*

```{r question 1}
data <- mtcars 

data <- data %>% rownames_to_column (var = "Model") 

head(data, n = 5)
```

***2.** Repurposing old code is not only the case for data mutations; we
can also use it for plots (or anything else). Additionally, in some
cases you may not even need to change your code at all if you’re
answering the same question or re-analyzing data with some additional
steps.*

***a.** For question 2, refer back to the previous code you wrote in
“Lab 2 – ggplot” to recreate this plot. Note, the below plot used the
theme_bw() function and you should not have to write any unique code to
recreate this plot (including title and axes names).*

***1.** Save this as plot_orig*

```{r question 2}
plot_orig <- ggplot(data = data, aes(x = wt, y = mpg)) + geom_point() +
  stat_smooth(method = "lm", se = FALSE) +
  ggtitle("Relationship of Car Weight to Miles per Gallon") + 
  labs(x = "Car weight (in tons)", y = "Miles per gallon", shape = "Transmission Type")+
  theme_bw()



plot_orig
```

***3.** Additionally, you may find that you can regularly use old code
as a “skeleton’ and then add to it to create something new. For example,
let’s say I recently found out that I would be presenting this plot to a
large lecture hall rooms with hundreds of rows and I wanted to make the
dots bigger so that people in the last rows would still be able to see
them. To do so, I could re-use my previous code while making one
change.*

***a.** For question 3, repurpose your old code to recreate the previous
plot, but change the size of the dots to be 4 instead of the default
(1).*

***i.** Save this as plot_sized*

```{r question 3}
plot_sized <- ggplot(data = data, aes(x = wt, y = mpg)) + 
  geom_point(aes(), size = 4) +
  stat_smooth(method = "lm", se = FALSE) +
  ggtitle("Relationship of Car Weight to Miles per Gallon") +
  labs(x = "Car weight (in tons)", y = "Miles per gallon", shape = "Transmission Type") +
  theme_bw()


plot_sized
```

***4.** Another big part of learning to program is learning how to
troubleshoot using error messages as guides. Building on our previous
example, someone in the audience loved my talk and now they want me to
deliver it to a group of car enthusiasts in a small conference room.
Because of the smaller venue, I no longer need the bigger dots, so we
can remove that from our code. However, I also know these car
enthusiasts care about whether the cars are automatic (car automatically
changes gears as needed) or manual (driver must change gears as the car
requires it). Accordingly, I planned to use the below code to change the
shape to reflect whether the car is manual or automatic (which is
captured in the “am” column of our datset, but I realized something
isn’t quite working when I run my planned code.*

***a.** Code*

***i.** Original code for Q1 with one change on the geom_point call*

***1.** Original code for geom_point: geom_point()*

***2.** Updated code for geom_point: geom_point(aes(shape = am))*

***b.** For question 4, read the error message and then build off our
previous lessons about using mutate to change a variable’s type in order
to allow the code to run. Tell me in a sentence or two how you thought
through what to do to resolve the error.*

2.  Save this as plot_shapes

```{r question 4 error code}

#This chunk of code produces the following error when attempting to manipulate the shape of the points in the scatter plot.  

#plot_shapes <- ggplot(data = data, aes(x = wt, y = mpg)) +
  #geom_point(aes(shape = am)) +
  #labs(x = "Car weight (in tons)", y = "Miles per gallon", shape = "Transmission Type") +
  #stat_smooth(method = "lm", se = FALSE) 

#plot_shapes #attempts to call scatter plot, but runs into an error. Most notable is error text "A continuous variable cannot be mapped to the shape aesthetic."

#Using 'str' we can look at the structure of the variables and see that geom_point is considering 'am' as a continuous variable when it should be a discrete variable.
#str(data$am)

#This section was commented out due to the chunk causing errors when knitting to a html_document.
```

```{r question 4 edit}
#This section of code is set to change the classification of variable 'am' from a continuous variable to a discrete variable
data <- data %>%
  mutate(am = as.factor((am))) 

#We then attempt to run the code again but with the new classification for 'am'
plot_shapes <- ggplot(data = data, aes(x = wt, y = mpg)) +
  geom_point(aes(shape = am)) +
  labs(x = "Car weight (in tons)", y = "Miles per gallon", shape = "Transmission Type") +
  scale_shape_discrete(labels = c("Automatic", "Manual")) +
  stat_smooth(method = "lm", se = FALSE) +
  ggtitle("Relationship of Car Weight to Miles per Gallon")+
  theme_bw()
 
plot_shapes 
```

***5.** One of the nice parts about coding is we can often take the
lessons we’ve learned from various other parts of code and put it
together.*

***a.** For question 5, change your geom_point call so it varies both
the shapes by am and sets the size of the points to 4.*

***b.** Save this as plot_final*

```{r question 5}
plot_final <- ggplot(data = data, aes(x = wt, y = mpg)) +
  geom_point(aes(shape = am), size = 8) +
  labs(x = "Car weight (in tons)", y = "Miles per gallon", shape = "Transmission Type") +
  scale_shape_discrete(labels = c("Automatic", "Manual")) +
  stat_smooth(method = "lm", se = FALSE) + 
  ggtitle("Relationship of Car Weight to Miles per Gallon") +
  theme_bw()
  
plot_final  



```

***6.** The car enthusiasts loved my last plot and want me to deliver it
another time at a national conference within a huge lecture hall. I have
a weird hunch that they’re going to want to know which cars correspond
to which dots.*

***a.** For question 6, use the ggplotly function to identify which car
is the rightmost dot in the previous plot. Report the following
information on that dot:*

***i.** The car model* ***ii.** Whether the car has an automatic or
manual transmission system* ***iii.** The reported weight of the car*
***iv.** The reported miles per gallon of the car*

```{r question 6}
plotly_plot <- ggplot(data = data, aes(x = wt, y = mpg, Model = Model)) +
  geom_point(aes(shape = am), size = 9) +
  scale_shape_discrete(labels = c("Automatic", "Manual")) +
  labs(x = "Car weight (in tons)", y = "Miles per gallon", shape = "Transmission Type") +
  stat_smooth(method = "lm", se = FALSE) + 
  ggtitle("Relationship of Car Weight to Miles per Gallon") 

ggplotly(plotly_plot)


```

*Car model: Lincoln Continental* *Transmission system: Automatic*
*Weight: 5.424 tons* *MPG: 10.4*

***7.** A budding data scientist sees my presentation and is wowed by my
final graph. They want to know how I iterated over time to create what
they called “The most beautiful graph about miles per gallon and car
weight that they’ve ever seen.”*

***a.** For question 7, use ggarrange to position all 4 plots within one
figure that has two columns, two rows, and legend set to “none”*

```{r question 7}
plot_orig <- plot_orig + ggtitle("")

plot_sized <- plot_sized + ggtitle("") 

plot_shapes <- plot_shapes + ggtitle("")

plot_final <- plot_final + ggtitle("")

ggarrange(plot_orig, plot_sized, plot_shapes, plot_final, nrow = 2, ncol = 2, legend = "none")
```

***8.** Our new figure looks mostly good, but the titles are far too
bunched. Since they’re so repetitive, we’re going to make a change to
the styling of our ggplots.*

***a.** For question 8, do the following:*

***i.** Remove the titles from the individual plots*

***ii.** Set the title of your ggarrange figure to “The relationship
between car weight and miles per gallon” and set it in a bold blue font
that has a small enough size that all words in your title can be read.*

```{r question 8}
ggarrange_q8 <- ggarrange(plot_orig, plot_sized, plot_shapes, plot_final, nrow = 2, ncol = 2, legend = FALSE) 

ggarrange_q8 <- annotate_figure(ggarrange_q8, top = text_grob("The relationship between car weight and miles per gallon", color = "blue", face = "bold"))

ggarrange_q8
```

***9.** While I like the 2nd and 3rd plots we made (plot_sized,
plot_shapes), I think that just showing the original plot (plot_orig)
and the final plot (plot_final) vertically would allow the student to
best see how far we came from our original plot until the final draft.*

***a.** For question 9, alter your ggarrange call so that you only have
the original and final plot stacked vertically on top of each other*

```{r question 9}
ggarrange_q9 <- ggarrange(plot_final, plot_orig, nrow = 2, ncol = 1, legend = FALSE)

ggarrange_q9 <- annotate_figure(ggarrange_q9, top = text_grob("The relationship between car weight and miles per gallon", color = "blue", face = "bold"))

ggarrange_q9

```

***10.** For question 10, alter your ggarrange call so that you only
have the original and final plot placed horizontally next to each other.
Then tell me which of the final 2 plots (this plot and the one created
in Q9) you would opt to use in a presentation within our class and why.*

```{r question 10}
ggarrange_q10 <- ggarrange(plot_orig, plot_final, nrow = 1, legend = FALSE)

ggarrange_q10 <- annotate_figure(ggarrange_q10, top = text_grob("The relationship between car weight and miles per gallon", color = "blue", face = "bold"))

ggarrange_q10

#I would use the plot from question 10. The plot from question 9 looks altered and may convey the wrong information simply due to the way the graphs look.


```
