---
title: 'Lab 3: dplyr'
author: "Isaias Perez"
date: "`r Sys.Date()`"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*#1. To begin, we are going to use the ‘nflfastR’ beginner’s guide to figure out how to load data from the 2022 NFL season into R.*

```{r Installing packages}
#install.packages("tidyverse")
#install.packages("nflfastR")
#install.packages("nflreader")
#install.packages("dplyr")
```

```{r loading installed packages}
library(tidyverse)
library(nflfastR)
library(nflreadr)
library(dplyr)
```

```{r adjusting scientific notation}
options(scipen = 9999)
```

```{r loading pre-cleaned data and assigning variable name}
data <- load_pbp(2022)
```

*\# For question 2, use the R Markdown environment to determine how many rows and columns this dataset includes. Additionally, tell me why we may consider using dplyr to reduce the size of our dataset if our intent is only to examine how Tampa Bay Buccaneer’s Quarterback Tom Brady played in 2022; within this response, be sure to mention some specific columns or rows that may be unnecessary.*

```{r evaluating rows & columns}
# dim allows us to view the dimensions of the data and to get a sense of the rows and columns associated with the dataset.
dim(data)
# str is short for structure; it is a more in-depth view of the what the data set is.
str(data[1:20])
```

*\# Within the data set, there are 50,147 rows & 372 columns. "Dim(data)" allows us to view this information. It is important to filter this dataset down as much as possible to be able to ask specific questions. In particularly, we are interested in evaluating Tampa Bay Buccaneer's quarterback Tom Brady performance during the 2022 session, and certain columns/ rows may not be necessary to view. For example, the following columns quarter_seconds_remaining, half_seconds_remaining, and game_seconds_remaining would not provide good insight as to how well Tom Brady played and could be an example of variables being filtered out.*

*\# 3. As previously noted, this dataset is HUGE. We can use dplyr to shorten it. To begin, lets filter our data so that we only have Buccaneers games from the regular season and plays where the ball was thrown that did not immediately follow a touchdown (these are the plays used to determine one’s ‘regular season stats’). We can filter for this information by using the following code:*

```{r filtering data}

# This chunk of code helps filter the data set and lowers the original 50,147 rows down to 1,321.

data <- data %>%    # This section is responsible for reassigning the filtered data set back to the variable name "data"
  filter(home_team == c("TB") | away_team == c("TB")) %>%  #This section is responsible for filtering the data set to only included homegames with TB and awaygames with TB.
  filter(season_type == c("REG")) %>%  # This section of code filters anything this is not apart of the regular football season.
  filter(pass_attempt == 1) %>% # This section is dedicated to retain data that only has pass plays.
  filter(two_point_attempt == 0) %>% # This section is filters out two point attempt plays.
  filter(sack == 0) # This section filters out plays that ended in a QB sack.
```

*#4. At this point, we have reduced our dataset down considerably as it pertains to rows. However, we still have a few hundred columns, many of which are irrelevant to our question regarding Buccaneers Quarterback Tom Brady. To make our dataset more manageable, we can use the below code:*

```{r filtering columns}
# The following line of code will help filter the data set further. It will do this by piping the data set through the select filter and keep only the variables in the parenthesis. This will lower the total column count from 372 down to 14.
data <- data%>%
  select(game_id, week, home_team, away_team, desc, pass_attempt, complete_pass, yards_gained, incomplete_pass, interception, sack, touchdown, passer_player_name, passing_yards)
```

*#5. At this point, we should only have data from passing plays in games when the Buccaneers were playing. Importantly, Tom Brady only played for the Buccaneers this season, which means all of his passes are included within this dataset. However, we missed a crucial piece of cleaning to this point. For question 5, tell me what (or who) else needs to be removed to ensure that we are only examining data pertaining to passing plays that involve Buccaneers Quarterback Tom Brady.*

```{r unique function to view all vlaues that are in a column}
unique(data$passer_player_name)

# everyone who is not Tom Brady will be need to be removed to understand data related to passing plays by the Buccaneers QB.

```

*\# For question 6, edit the previous lines of code that were used to filter data and only keep plays that involve Tom Brady. To check that your code worked, use the ‘unique(data\$passer_player_name)’ and tell me what changed and how you know your code accomplished what you set out to do.*

```{r additonal filter to manage QBs}
data <- data %>%
  filter(passer_player_name == c("T.Brady")) # This line of code passes the data set through an additional filter and keeps only rows with Tom Brady's name within the Passer_player_name column.


unique(data$passer_player_name) # This sections is meant to be a second check to make sure that the filter was applied.
```

*#For question 7, do the following: 1) Add hash tags to each line of code and tell me what it does ‘in human terms’. 2) Write a one sentence summary preceding the chunk that tells me what it does.*

```{r group_by & summarize functions}
data <- data %>% replace(is.na(.),0) # this will turn any blank cells into 0's.

week_passing <- data %>% #creating a new variable.
  group_by(week) %>% # group_by function will separate groups based on the variable of week.
  summarise(           #summaries function will sum together the following variables of pass_attempts, complete_pass, passing_yards, and touchdowns. 
    attempts = sum(pass_attempt),
    completions = sum(complete_pass),
    yards = sum(passing_yards),
    touchdowns = sum(touchdown)
    )%>%
  arrange(week)       # this line will arrange the new data set by week.
week_passing         # this line will simply display the dataset.
```

*#For question 8, I want to know how many yards Tom Brady gained on average every time we threw the ball in each week. Additionally, we may want to how what percentage of passes Tom Brady threw that were caught by a player on his team for a pass completion.*

```{r yards per attempt & completion percentage}
#The bottom chunk's are used to create new columns within the data set of week_passing. 
week_passing <- 
  week_passing %>%
  mutate("YPA" = yards/attempts)

week_passing <- 
  week_passing %>%
  mutate("CP" = completions/ attempts)

head(week_passing) #This allows to briefly view what we changed.
```

*\# For question 9, add hashtags to each line of code and tell me what it does ‘in human terms’ and tell me if you see any trend*

```{r visualizing}
ggplot(                       #ggplot is the package which allows for graph assembly.
  data = week_passing,        #This line is shows R which data set to access.
  aes(x = as.factor(week), y = attempts) #aes is needed to assign variables to the x and y axis.
)+
  geom_bar(stat = "identity")+  #this section determines the type of graph being used and specifies bar height.
  theme_bw() +        # this apllies a black and white theme.
  ggtitle("Tom Brady's pass attempts by week")+   #this line assigns a title.
  labs(y = "pass attempts", x = "Week of NFL season") #assigns labels for each axis.

#From the graph alone, it does not appear to have a distinctive relationship.

```

*\# For question 10, I want you to create a second visualization using any other variable or variables within the week_passing dataset (can be a box plot, a box and whisker plot, or a scatter plot) and tell me something else about Tom Brady’s season.*

```{r}
ggplot(week_passing, aes(x = week, y = yards, color = factor(week))) +
  geom_point() +
  theme_bw() +
  theme(legend.key.size = unit(0.3, "cm"))+
  scale_x_continuous(name = "Reg. Season Weeks", breaks = c(2,4,6,8,10,12,14,16,18))+
  scale_y_continuous(name = "Yards Passed Per Game", breaks = c(50,100,150,200,250,300,350,400))
  

#This additional graph has weeks plotted on the x-axis and yards passed on the y-axis. We can see that spread of yards per game was moderately consistent between weeks 3 and 16. For weeks 1, 2, 17, and 18 the variance between the points appeared much greater.


```
