---
title: "Lab 10 Writing paper in R"
author: "Isaias Perez"
date: "`r Sys.Date()`"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r installing packages, echo=FALSE}
#install.packages("here")
#install.packages("tidyverse")
#install.packages("psych")
#install.packages("papaja")
```

```{r library, echo=FALSE}
library(here)
library(tidyverse)
library(psych)
library(dplyr)
library(papaja)

options(scipen = 999)
```

***1.** Once again, I want to give us practice using the “here()” call.* ***a.** For question 1, please import the dataset ‘pizzaplace.csv’ from Canvas using the here function and save it as “data.” Also use the below mutate call to make one of the column a bit more interpretable*

```{r questions 1}
data <- read.csv(here("pizzaplace.csv")) #calls csv file to global enviroment.

#The bottom code recodes variable names.
data <- data %>% mutate(size = recode(size, 
                                      "S" = "small",
                                      "M" = "medium",
                                      "L" = "large",
                                      "XL" = "extra large",
                                      "XXL" = "extra large"))


```

***2.** Next, I want to start with simply putting text into the document.*

***a.** For question 2, outside of a code chunk, write the follwing sentence: i. The first pizza sold in 2015 was a TYPE pizza.*

***b.** In place of the all caps words, manually plug in data from the first row of the data within the dataset just like you would if I asked you to do this a Word document*

The first pizza sold in 2015 was a hawaiian pizza.

***3.** Next, I’d like to practice automating that code.* ***a.** For question 3, allow R to automate the TYPE of the pizza for you with an in-line chunk. Use the below code*

```{r question 3}
data$type[1] #This line access the dataset and reveals the type of pizze located on row 1.
```

#question 3 sentence construction The first pizza sold in 2015 was a `r data$type[1]` pizza.

\*\*\*\*a.\*\* For question 4, follow the framework laid out in question 3 to add to your sentence by plugging in the values for all of the ‘all caps’ words: i. The first pizza sold in 2015 was a SIZE TYPE pizza that sold for PRICE.\*

```{r question 4}

#These lines displays the size and price of the pizza of row 1.
data$size[1]
data$price[1]

```

# question 4 sentence construction.

The first pizza sold in 2015 was a `r data$size[1]`, `r data$type[1]` pizza that sold for `r data$price[1]`

***a**. For question 5, let’s assume the price column doesn’t account for tax (let’s assume 10% just for easy math). Make a new column called full_price within the dataset (think of what call we need for a new column) that is the cost of the pizza plus tax and then automate the below sentence:i. The first pizza sold in 2015 was a SIZE TYPE pizza that sold for PRICE. The customer paid FULL_PRICE with tax.*

```{r question 5}
#This line creates two new columns within the data set. The new tax column takes all observations of the column price and multiplies it by .10. The new full_price column takes the price column and the new tax column and add them both together.
data <- data %>% mutate(tax = price * .10,
                        full_price = price + tax)

data$full_price[1]
```

# question 5 sentence construction.

The first pizza sold in 2015 was a `r data$size[1]` , `r data$type[1]` pizza that sold for `r data$price[1]`. The customer paid `r data$full_price[1]` with tax.

***6.**use the print_num() function covered in class and set digits to 2. Then recreate the sentence so that FULL_PRICE only has 2 digits.*

```{r question 6}
#Same as code for question 5.
data <- data %>% mutate(tax = price * .10,
                        full_price = price + tax)

#This line applies a deicmal limit of 2.
data$full_price <- print_num(data$full_price, digits = 2)

head(data$full_price)
```

# question 6 sentence construction.

The first pizza sold in 2015 was a `r data$size[1]` , `r data$type[1]` pizza that sold for `r data$price[1]`. The customer paid `r data$full_price[1]` with tax.

***7.** recreate the above sentence, but instead of using the value in full_price for the FULL_PRICE part of the sentence, use the value for price and multiply it by 1.1 within the in-line chunk for FULL_PRICE. Once again round your digits to 2.*

# question 7 sentence construction.

The first pizza sold in 2015 was a `r data$size[1]` , `r data$type[1]` pizza that sold for `r data$price[1]`. The customer paid `r data$price[1]* 1.1` with tax.

***8** hashtag each line of the code that I’ve written with what it does. At the bottom (xvii and xviii) I wrote the function to get the data from row 2. After doing so, use the str() call to identify where each aspect of data you’ll need to recreate the last version of the sentence (shown below) is located within your output*

```{r question 8 function, echo=FALSE}
data$price <- as.numeric(data$price)

#This function is used to automate the retrieval of key information such as pizza type, size, price, and its full price.
pizza_info <- function(rownum, data){
  pizza <- data %>% dplyr::filter(X == rownum) %>% as.data.frame() 
  type <- pizza %>% dplyr::select(type) %>% as.vector()
  size <- pizza %>% dplyr::select(size) %>% as.vector()
  price <- pizza %>% dplyr::select(price) %>% as.vector()
  pizza <- pizza %>% dplyr::mutate(full_price = price * 1.1) %>% as.data.frame()
  full_price <- pizza %>% dplyr::select(full_price) %>% as.vector()
  pizza_list <- list("type" = type, 
	                     "size" = size, 
	                     "price" = price,
	                     "full_price" = full_price) 
  return(pizza_list)
  }
```

```{r question 8 continued}
#Using the new function. We can find the type, size, price, and full price of the pizza on row 2 of the dataset.
pizza_sentence_info <- pizza_info(2, data)
pizza_sentence_info

str(pizza_sentence_info)
```

# question 8 sentence construction.

The first pizza sold in 2015 was a `r pizza_sentence_info$size[1]`, `r pizza_sentence_info$type[1]` pizza that sold for `r pizza_sentence_info$price[1]`. The customer paid `r pizza_sentence_info$full_price[1]` with tax.

***9.** Now that we’ve created a function to store our data in a consistent and easy to access place, we can write as many sentence as we want with ease.* ***a.** For question 9, recreate the sentence for the follow row numbers: 5, 42, 398, 1776, 10402, 49000*

```{r question 9 row # 5}
pizza_sentence_5 <- pizza_info(5, data)
pizza_sentence_5

```

The first pizza sold in 2015 was a `r pizza_sentence_5$size[1]`, `r pizza_sentence_5$type[1]` pizza that sold for `r pizza_sentence_5$price[1]`. The customer paid `r pizza_sentence_5$full_price[1]` with tax.

```{r question 9 row # 42}
pizza_sentence_42 <- pizza_info(42, data)
pizza_sentence_42

```

The first pizza sold in 2015 was a `r pizza_sentence_42$size[1]`, `r pizza_sentence_42$type[1]` pizza that sold for `r pizza_sentence_42$price[1]`. The customer paid `r pizza_sentence_42$full_price[1]` with tax.

```{r question 9 row # 398}
pizza_sentence_398 <- pizza_info(398, data)
pizza_sentence_398
```

The first pizza sold in 2015 was a `r pizza_sentence_398$size[1]`, `r pizza_sentence_398$type[1]` pizza that sold for `r pizza_sentence_398$price[1]`. The customer paid `r pizza_sentence_398$full_price[1]` with tax.

```{r question 9 row # 1776}
pizza_sentence_1776 <- pizza_info(1776, data)
pizza_sentence_1776
```

The first pizza sold in 2015 was a `r pizza_sentence_1776$size[1]`, `r pizza_sentence_1776$type[1]` pizza that sold for `r pizza_sentence_1776$price[1]`. The customer paid `r pizza_sentence_1776$full_price[1]` with tax.

```{r question 9 row # 10402}
pizza_sentence_10402 <- pizza_info(10402, data)
pizza_sentence_10402
```

The first pizza sold in 2015 was a `r pizza_sentence_info$size[1]`, `r pizza_sentence_info$type[1]` pizza that sold for `r pizza_sentence_info$price[1]`. The customer paid `r pizza_sentence_info$full_price[1]` with tax.

```{r question 9 row # 49000}
pizza_sentence_49000 <- pizza_info(49000, data)
pizza_sentence_49000


```

The first pizza sold in 2015 was a `r pizza_sentence_info$size[1]`, `r pizza_sentence_49000$type[1]` pizza that sold for `r pizza_sentence_49000$price[1]`. The customer paid `r pizza_sentence_49000$full_price[1]` with tax.

***10.** We can also pull data from tables we make in addition to the raw data that we import.*

***a.** For question 10, use the below code to create descriptive tables on the price for all pizza sizes. Use the str() function to find the location of the mean and standard deviation of the price of pizza for each size of pizza. Then report the mean and standard deviation of the price for each pizza size within an automated paragraph.*

```{r question 10}
#This bottom line creates four dataframes separated by pizza size to show descriptive statistics of each one.
descriptives <- psych::describeBy(data, group = "size")


# 1 = extra large
XL_mean <- print_num(descriptives$"1"$mean[8], digits = 2)
XL_sd <- print_num(descriptives$"1"$sd[8], digits = 2)

#2 = large 
large_mean <- print_num(descriptives$"2"$mean[8], digits = 2)
large_sd <- print_num(descriptives$"2"$sd[8], digits = 2)

#3 = medium
medium_mean <- print_num(descriptives$"3"$mean[8], digits = 2)
medium_sd <- print_num(descriptives$"3"$sd[8], digits = 2)

# 4 = small
small_mean <- print_num(descriptives$"4"$mean[8], digits = 2)
small_sd <-print_num(descriptives$"4"$sd[8], digits = 2)
```

# question 10 sentence construction

For pizza's sold in 2015, the pizza size greatly influenced its sale price. For instance, extra large pizza's sold at an avg price of `r XL_mean` with a sd of `r XL_sd`. Large pizza's sold for an avg price of `r large_mean` with a sd of `r large_sd`. Medium size pizza's sold for an avg price of `r medium_mean` with a sd of `r medium_sd`. Finally, small pizza's sold for an avg price of `r small_mean` with a sd of `r small_sd`.
